---
title: "Code Portfolio"
author: "Jiayin Liang - 4/14/2019"
date: "April 14, 2019"
output: html_document
---

```{r}
```

## R Markdown

```{r}
```

## Including Plots

```{r}
```

1. Data Import

(1) Reads comma delimited files

```{r}
read_csv()
```
(2) Reads semicolon separated files

```{r}
read_csv2()
```
(3) Reads files with tab delimited

```{r}
read_tsv()
```
(4) Reads in files with any delimiter

```{r}
read_delim()
```
(5) Reads files with fixed width

```{r}
read_fwf()
```
(6) Reads a file into data frame in table format

```{r}
read_table()
```
(7) Reads Apache style log files

```{r}
read_log()
```

2. Data Wrangling

(1) Select columns of a data frame

```{r}
select()
```
(2) Choose rows based on specific criteria

```{r}
filter()
```
(3) Select certain rows by row number

```{r}
slice()
```
(4) Sort rows by variables

```{r}
arrange()
```
(5) Create new columns based on the values in existing columns

```{r}
mutate()
```
(6) Splits the data into groups

```{r}
group_by()
```
(7) Collapse each group into a single-row summary

```{r}
summarize()
```
(8) Count the number of observations found for each factor or combination of factors

```{r}
tally()
```

3. Data Visualization

(1) Create bar chart to show a categorical comparison between two or more variables

```{r}
barplot()
```
(2) Histogram: to show the probability distribution of a continuous variable

```{r}
hist()
```
(3) Scatter plot: tracks two continuous variables over the course of time

```{r}
plot()
```
(4) Box Plot: provide visualization for both categorical and continuous variable data

```{r}
boxplot()
```
(5) Creates a coordinate system that can add layers to

```{r}
ggplot()
```
(6) Adds a layer of points to the plot

```{r}
geom_point()
```
(7) Split your plot into facets, subplots that each display one subset of the data. Facet plot by a single variable:

```{r}
facet_wrap()
```
(8) Facet the plot on the combination of two variables

```{r}
facet_grid()
```
(9) To set the linetype of a line in the plot

```{r}
geom_smooth()
```

4. Data Analysis

(1) Compute a distance matrix between the rows of a data matrix

```{r}
get_dist()
```
(2) Visualize a distance matrix

```{r}
fviz_dist()
```
(3) Compute k-means

```{r}
kmeans()
```
(4) To illustrate the clusters

```{r}
fviz_cluster()
```
(5) Elbow method: to find the appropriate number of clusters in a dataset

```{r}
fviz_nbclust(df, kmeans, method = "wss")
```
(6) Silhouette Method: to determines how well each object lies within its cluster, measures the quality of a clustering

```{r}
fviz_nbclust(df, kmeans, method = "silhouette")
```
(7) To calculate the gap statistic and standard error

```{r}
clusGap()
```
(8) To visualize the result of gap statistic method

```{r}
fviz_gap_stat()
```

